A Spectral Two-Moment Model \\ for Neutrino-Matter Interactions

### 2019-06-14 Documents/M1/collisionsM1/collisionsM1.tex

\newcommand{\IN}{\mbox{\tiny In}}
\newcommand{\OUT}{\mbox{\tiny Out}}
\newcommand{\NNS}{\mbox{\tiny NNS}} % Neutrino-Nucleon/Nuclei Scattering
\newcommand{\NES}{\mbox{\tiny NES}} % Neutrino-Electron Scattering
\newcommand{\PROD}{\mbox{\tiny Pr}}
\newcommand{\ANN}{\mbox{\tiny An}}
\newcommand{\SC}{\mbox{\tiny Sc}}      % Scattering
\newcommand{\TP}{\mbox{\tiny TP}}      % Pair-Processes

@@

@@

>>>>

\newcommand{\EC}{\mbox{\tiny{\sc Ec}}}
\newcommand{\IN}{\mbox{\tiny{\sc In}}}
\newcommand{\OUT}{\mbox{\tiny{\sc Out}}}
\newcommand{\NNS}{\mbox{\tiny NNS}} % Neutrino-Nucleon/Nuclei Scattering
\newcommand{\NES}{\mbox{\tiny{\sc Nes}}} % Neutrino-Electron Scattering
\newcommand{\PROD}{\mbox{\tiny{\sc Pr}}}
\newcommand{\ANN}{\mbox{\tiny{\sc An}}}
\newcommand{\SC}{\mbox{\tiny{\sc Sc}}}      % Scattering
\newcommand{\TP}{\mbox{\tiny{\sc Tp}}}      % Pair-Processes

@@

\section{Implicit Solvers}

We discuss the implicit solver in the context of a system of electron neutrinos and antineutrinos.  
We let $\mathcal{J}_{\nu_{e}}\to\mathcal{J}$ and $\mathcal{J}_{\bar{\nu}_{e}}\to\bar{\mathcal{J}}$.  
The electron fraction is updated from
\begin{equation}
  \f{\rho}{m_{b}}\big(Y_{e}^{n+1}-Y_{e}^{n}\big) + \big[\,\big(J^{n+1}-\bar{J}^{n+1}\big)-\big(J^{n}-\bar{J}^{n}\big)\,\big] = 0,
  \label{eq:electronFractionContinuous}
\end{equation}
where
\begin{equation}
  \big\{\,J,\bar{J}\,\big\}
  =\sum_{q=1}^{N_{\varepsilon}}\hat{w}_{q}^{(2)}\,\big\{\,\mathcal{J}_{q},\bar{\mathcal{J}}_{q}\,\big\}
  \quad\text{with}\quad
  \hat{w}_{q}^{(2)}=\f{4\pi}{h^{3}}\,w_{q}^{(2)},
\end{equation}
where $w_{q}^{(2)}$ are quadrature weights for energy integrals with respect to $dV_{\varepsilon}=\varepsilon^{2}\,d\varepsilon$.  
Defining $S_{Y_{e}}\equiv \rho\,Y_{e}^{n}/m_{b}$ and $U_{Y_{e}} \equiv Y_{e}/Y_{e}^{n}$, we can write Eq.~\eqref{eq:electronFractionContinuous} in discrete form as
\begin{equation}
  F_{Y_{e}}(\vect{U}^{n+1})\equiv
  U_{Y_{e}}^{n+1} + \sum_{q}\hat{w}_{q}^{(2)}\big(\,\mathcal{J}_{q}^{n+1}-\bar{\mathcal{J}}_{q}^{n+1}\,\big)/S_{Y_{e}} 
  - \big\{\,1+\sum_{q}\hat{w}_{q}^{(2)}\big(\,\mathcal{J}_{q}^{n}-\bar{\mathcal{J}}_{q}^{n}\,\big)/S_{Y_{e}}\,\big\} = 0.
  \label{eq:electronFractionDiscrete}
\end{equation}
Here we have defined the solution vector $\vect{U}=\big(Y_{e},\epsilon,\vect{\mathcal{J}},\vect{\bar{\mathcal{J}}}\big)^{T}$, where $\vect{\mathcal{J}}=\big(\mathcal{J}_{1},\ldots,\mathcal{J}_{N_{\varepsilon}}\big)^{T}$ and $\vect{\bar{\mathcal{J}}}=\big(\bar{\mathcal{J}}_{1},\ldots,\bar{\mathcal{J}}_{N_{\varepsilon}}\big)^{T}$.  
We also define the vectors $\vect{u}=\big(Y_{e},\epsilon\big)^{T}$ and $\vect{\mathcal{U}}=\big(\vect{\mathcal{J}},\vect{\bar{\mathcal{J}}}\big)^{T}$.

The specific internal energy is updated from
\begin{equation}
  \rho\,\big(\epsilon^{n+1}-\epsilon^{n}\big)
  +\big[\,\big(E^{n+1}+\bar{E}^{n+1}\big)-\big(E^{n}+\bar{E}^{n}\big)\,\big] = 0,
  \label{eq:internalEnergyContinuous}
\end{equation}
where
\begin{equation}
  \big\{\,E,\bar{E}\,\big\}
  =\sum_{q=1}^{N_{\varepsilon}}\hat{w}_{q}^{(3)}\,\big\{\,\mathcal{J}_{q},\bar{\mathcal{J}}_{q}\,\big\}
  \quad\text{with}\quad
  \hat{w}_{q}^{(3)}=\f{4\pi}{h^{3}}\,w_{q}^{(3)},
\end{equation}
where $w_{q}^{(3)}$ are quadrature weights for energy integrals with respect to $\varepsilon\,dV_{\varepsilon}=\varepsilon^{3}\,d\varepsilon$.  
Defining $S_{\epsilon}\equiv\rho\epsilon^{n}$ and $U_{\epsilon}\equiv\epsilon/\epsilon^{n}$, we can write Eq.~\eqref{eq:internalEnergyContinuous} as
\begin{equation}
  F_{\epsilon}(\vect{U}^{n+1})\equiv
  U_{\epsilon}^{n+1} + \sum_{q}\hat{w}_{q}^{(3)}\big(\,\mathcal{J}_{q}^{n+1}+\bar{\mathcal{J}}_{q}^{n+1}\,\big)/S_{\epsilon}
  -\big\{\,1+\sum_{q}\hat{w}_{q}^{(3)}\big(\,\mathcal{J}_{q}^{n}+\bar{\mathcal{J}}_{q}^{n}\,\big)/S_{\epsilon}\,\big\} = 0.  
  \label{eq:internalEnergyDiscrete}
\end{equation}

The neutrino density is updated from
\begin{align}
  \f{\big(\mathcal{J}_{i}^{n+1}-\mathcal{J}_{i}^{n}\big)}{\dt}
  &=\big[\chi_{\EC}\big]_{i}^{n}\,\big(\mathcal{J}_{0,i}^{n+1}-\mathcal{J}_{i}^{n+1}\big) \nonumber \\
  &\hspace{12pt}
  +\sum_{q}w_{q}^{(2)}\,\big[\Phi_{\NES,0}^{\In}\big]_{qi}^{n+1}\,\mathcal{J}_{q}^{n+1} \nonumber \\
  &\hspace{12pt}
  -\mathcal{J}_{i}^{n+1}\sum_{q}w_{q}^{(2)}\,\big\{\,\big[\Phi_{\NES,0}^{\In}\big]_{qi}^{n+1}\,\mathcal{J}_{q}^{n+1}+\big[\Phi_{\NES,0}^{\Out}\big]_{qi}^{n+1}\,\big(1-\mathcal{J}_{q}^{n+1}\big)\,\big\} \nonumber \\
  &\hspace{12pt}
  + \sum_{q}w_{q}^{(2)}\,\big[\Phi_{\TP,0}^{\PROD}\big]_{qi}^{n+1}\,\big(1-\bar{\mathcal{J}}_{q}^{n+1}\big) \nonumber \\
  &\hspace{12pt}
  -\mathcal{J}_{i}^{n+1}\sum_{q}w_{q}^{(2)}\big\{\,\big[\Phi_{\TP,0}^{\PROD}\big]_{qi}^{n+1}\,\big(1-\bar{\mathcal{J}}_{q}^{n+1}\big)+\big[\Phi_{\TP,0}^{\ANN}\big]_{qi}^{n+1}\,\bar{\mathcal{J}}_{q}^{n+1}\,\big\} \nonumber \\
  &=\big(\big[\eta_{\EC}\big]_{i}^{n+1}+\big[\eta_{\NES}\big]_{i}^{n+1}+\big[\eta_{\TP}\big]_{i}^{n+1}\big) \nonumber \\
  &\hspace{12pt}
  -\big(\big[\chi_{\EC}\big]_{i}^{n}+\big[\chi_{\NES}\big]_{i}^{n+1}+\big[\chi_{\TP}\big]_{i}^{n+1}\big)\,\mathcal{J}_{i}^{n+1} \nonumber \\
  &=\big[\eta\big]_{i}^{n+1} - \big[\chi\big]_{i}^{n+1}\,\mathcal{J}_{i}^{n+1},
  \label{eq:neutrinoDiscrete}
\end{align}
where we have defined
\begin{align}
  \big[\eta_{\EC}\big]_{i}^{n+1}
  &=\big[\chi_{\EC}\big]_{i}^{n}\,\mathcal{J}_{0,i}^{n+1}, \label{eq:ECemissivityNeutrino} \\
  \big[\eta_{\NES}\big]_{i}^{n+1}
  &=\sum_{q}w_{q}^{(2)}\,\big[\Phi_{\NES,0}^{\In}\big]_{qi}^{n+1}\,\mathcal{J}_{q}^{n+1}, \label{eq:NESemissivityNeutrino} \\
  \big[\eta_{\TP}\big]_{i}^{n+1}
  &=\sum_{q}w_{q}^{(2)}\,\big[\Phi_{\TP,0}^{\PROD}\big]_{qi}^{n+1}\,\big(1-\bar{\mathcal{J}}_{q}^{n+1}\big), \label{eq:PAIRemissivityNeutrino} \\
  \big[\chi_{\NES}\big]_{i}^{n+1}
  &=\sum_{q}w_{q}^{(2)}\,\big\{\,\big[\Phi_{\NES,0}^{\In}\big]_{qi}^{n+1}\,\mathcal{J}_{q}^{n+1}+\big[\Phi_{\NES,0}^{\Out}\big]_{qi}^{n+1}\,\big(1-\mathcal{J}_{q}^{n+1}\big)\,\big\} \\
  &=\big[\eta_{\NES}\big]_{i}^{n+1}+\sum_{q}w_{q}^{(2)}\,\big[\Phi_{\NES,0}^{\Out}\big]_{qi}^{n+1}\,\big(1-\mathcal{J}_{q}^{n+1}\big), \label{eq:NESopacityNeutrino} \\
  \big[\chi_{\TP}\big]_{i}^{n+1}
  &=\sum_{q}w_{q}^{(2)}\big\{\,\big[\Phi_{\TP,0}^{\PROD}\big]_{qi}^{n+1}\,\big(1-\bar{\mathcal{J}}_{q}^{n+1}\big)+\big[\Phi_{\TP,0}^{\ANN}\big]_{qi}^{n+1}\,\bar{\mathcal{J}}_{q}^{n+1}\,\big\} \\
  &=\big[\eta_{\TP}\big]_{i}^{n+1}+\sum_{q}w_{q}^{(2)}\,\big[\Phi_{\TP,0}^{\ANN}\big]_{qi}^{n+1}\,\bar{\mathcal{J}}_{q}^{n+1}, \label{eq:PAIRopacityNeutrino} \\
  \big[\eta\big]_{i}^{n+1}
  &=\big[\eta_{\EC}\big]_{i}^{n+1}+\big[\eta_{\NES}\big]_{i}^{n+1}+\big[\eta_{\TP}\big]_{i}^{n+1}, \label{eq:TOTALemissivityNeutrino} \\
  \big[\chi\big]_{i}^{n+1}
  &=\big[\chi_{\EC}\big]_{i}^{n}+\big[\chi_{\NES}\big]_{i}^{n+1}+\big[\chi_{\TP}\big]_{i}^{n+1}.  \label{eq:TOTALopacityNeutrino}
\end{align}
We rewrite Eq.~\eqref{eq:neutrinoDiscrete} and define
\begin{equation}
  \mathcal{F}_{i}(\vect{U}^{n+1})
  \equiv\big(\,1+\dt\,\big[\chi\big]_{i}^{n+1}\,\big)\,\mathcal{J}_{i}^{n+1}-\big(\,\mathcal{J}_{i}^{n}+\dt\,\big[\eta\big]_{i}^{n+1}\,\big)=0.
  \label{eq:neutrinoDiscreteResidualForm}
\end{equation}
Similarly, the antineutrino density is updated from
\begin{align}
  \f{\big(\bar{\mathcal{J}}_{i}^{n+1}-\bar{\mathcal{J}}_{i}^{n}\big)}{\dt}
  &=\big[\bar{\chi}_{\EC}\big]_{i}^{n}\,\big(\bar{\mathcal{J}}_{0,i}^{n+1}-\bar{\mathcal{J}}_{i}^{n+1}\big) \nonumber \\
  &\hspace{12pt}
  +\sum_{q}w_{q}^{(2)}\,\big[\bar{\Phi}_{\NES,0}^{\In}\big]_{qi}^{n+1}\,\bar{\mathcal{J}}_{q}^{n+1} \nonumber \\
  &\hspace{12pt}
  -\bar{\mathcal{J}}_{i}^{n+1}\sum_{q}w_{q}^{(2)}\,\big\{\,\big[\bar{\Phi}_{\NES,0}^{\In}\big]_{qi}^{n+1}\,\bar{\mathcal{J}}_{q}^{n+1}+\big[\bar{\Phi}_{\NES,0}^{\Out}\big]_{qi}^{n+1}\,\big(1-\bar{\mathcal{J}}_{q}^{n+1}\big)\,\big\} \nonumber \\
  &\hspace{12pt}
  + \sum_{q}w_{q}^{(2)}\,\big[\bar{\Phi}_{\TP,0}^{\PROD}\big]_{qi}^{n+1}\,\big(1-\mathcal{J}_{q}^{n+1}\big) \nonumber \\
  &\hspace{12pt}
  -\bar{\mathcal{J}}_{i}^{n+1}\sum_{q}w_{q}^{(2)}\big\{\,\big[\bar{\Phi}_{\TP,0}^{\PROD}\big]_{qi}^{n+1}\,\big(1-\mathcal{J}_{q}^{n+1}\big)+\big[\bar{\Phi}_{\TP,0}^{\ANN}\big]_{qi}^{n+1}\,\mathcal{J}_{q}^{n+1}\,\big\} \nonumber \\
  &=\big(\,\big[\bar{\eta}_{\EC}\big]_{i}^{n+1}+\big[\bar{\eta}_{\NES}\big]_{i}^{n+1}+\big[\bar{\eta}_{\TP}\big]_{i}^{n+1}\,\big) \nonumber \\
  &\hspace{12pt}
  -\big(\,\big[\bar{\chi}_{\EC}\big]_{i}^{n}+\big[\bar{\chi}_{\NES}\big]_{i}^{n+1}+\big[\bar{\chi}_{\TP}\big]_{i}^{n+1}\,\big)\,\bar{\mathcal{J}}_{i}^{n+1} \nonumber \\
  &=\big[\bar{\eta}\big]_{i}^{n+1} - \big[\bar{\chi}\big]_{i}^{n+1}\,\bar{\mathcal{J}}_{i}^{n+1},
  \label{eq:antineutrinoDiscrete}
\end{align}
where we have defined
\begin{align}
  \big[\bar{\eta}_{\EC}\big]_{i}^{n+1}
  &=\big[\bar{\chi}_{\EC}\big]_{i}^{n}\,\bar{\mathcal{J}}_{0,i}^{n+1}, \label{eq:ECemissivityAntineutrino} \\
  \big[\bar{\eta}_{\NES}\big]_{i}^{n+1}
  &=\sum_{q}w_{q}^{(2)}\,\big[\bar{\Phi}_{\NES,0}^{\In}\big]_{qi}^{n+1}\,\bar{\mathcal{J}}_{q}^{n+1}, \label{eq:NESemissivityAntineutrino} \\
  \big[\bar{\eta}_{\TP}\big]_{i}^{n+1}
  &=\sum_{q}w_{q}^{(2)}\,\big[\bar{\Phi}_{\TP,0}^{\PROD}\big]_{qi}^{n+1}\,\big(1-\mathcal{J}_{q}^{n+1}\big), \label{eq:PAIRemissivityAntineutrino} \\
  \big[\bar{\chi}_{\NES}\big]_{i}^{n+1}
  &=\sum_{q}w_{q}^{(2)}\,\big\{\,\big[\bar{\Phi}_{\NES,0}^{\In}\big]_{qi}^{n+1}\,\bar{\mathcal{J}}_{q}^{n+1}+\big[\bar{\Phi}_{\NES,0}^{\Out}\big]_{qi}^{n+1}\,\big(1-\bar{\mathcal{J}}_{q}^{n+1}\big)\,\big\} \\
  &=\big[\bar{\eta}_{\NES}\big]_{i}^{n+1}+\sum_{q}w_{q}^{(2)}\,\big[\bar{\Phi}_{\NES,0}^{\Out}\big]_{qi}^{n+1}\,\big(1-\bar{\mathcal{J}}_{q}^{n+1}\big), \label{eq:NESopacityAntineutrino} \\
  \big[\bar{\chi}_{\TP}\big]_{i}^{n+1}
  &=\sum_{q}w_{q}^{(2)}\big\{\,\big[\bar{\Phi}_{\TP,0}^{\PROD}\big]_{qi}^{n+1}\,\big(1-\mathcal{J}_{q}^{n+1}\big)+\big[\bar{\Phi}_{\TP,0}^{\ANN}\big]_{qi}^{n+1}\,\mathcal{J}_{q}^{n+1}\,\big\} \\
  &=\big[\bar{\eta}_{\TP}\big]_{i}^{n+1}+\sum_{q}w_{q}^{(2)}\,\big[\bar{\Phi}_{\TP,0}^{\ANN}\big]_{qi}^{n+1}\,\mathcal{J}_{q}^{n+1}, \label{eq:PAIRopacityAntineutrino} \\
  \big[\bar{\eta}\big]_{i}^{n+1}
  &=\big[\bar{\eta}_{\EC}\big]_{i}^{n+1}+\big[\bar{\eta}_{\NES}\big]_{i}^{n+1}+\big[\bar{\eta}_{\TP}\big]_{i}^{n+1}, \label{eq:TOTALemissivityAntineutrino} \\
  \big[\bar{\chi}\big]_{i}^{n+1}
  &=\big[\bar{\chi}_{\EC}\big]_{i}^{n}+\big[\bar{\chi}_{\NES}\big]_{i}^{n+1}+\big[\bar{\chi}_{\TP}\big]_{i}^{n+1}. \label{eq:TOTALopacityAntineutrino}
\end{align}
We rewrite Eq.~\eqref{eq:antineutrinoDiscrete} and define
\begin{equation}
  \bar{\mathcal{F}}_{i}(\vect{U}^{n+1})\equiv
  \big(\,1+\dt\,\big[\bar{\chi}\big]_{i}^{n+1}\,\big)\,\bar{\mathcal{J}}_{i}^{n+1}-\big(\,\bar{\mathcal{J}}_{i}^{n}+\dt\,\big[\bar{\eta}\big]_{i}^{n+1}\,\big)=0.
  \label{eq:antineutrinoDiscreteResidualForm}
\end{equation}

Next, we define
\begin{align}
  \vect{\mathcal{F}}(\vect{U})=\big(\,\mathcal{F}_{1}(\vect{U}),\ldots,\mathcal{F}_{N_{\varepsilon}}(\vect{U})\,\big)^{T}, \\
  \vect{\bar{\mathcal{F}}}(\vect{U})=\big(\,\bar{\mathcal{F}}_{1}(\vect{U}),\ldots,\bar{\mathcal{F}}_{N_{\varepsilon}}(\vect{U})\,\big)^{T}
\end{align}
and
\begin{equation}
  \vect{F}(\vect{U})=\big(F_{Y_{e}}(\vect{U}),F_{\epsilon}(\vect{U}),\vect{\mathcal{F}}(\vect{U}),\vect{\bar{\mathcal{F}}}(\vect{U})\big)^{T},
\end{equation}
so that we can write the coupled nonlinear system consisting of Eqs.~\eqref{eq:electronFractionDiscrete}, \eqref{eq:internalEnergyDiscrete}, \eqref{eq:neutrinoDiscreteResidualForm}, and \eqref{eq:antineutrinoDiscreteResidualForm} in the compact form
\begin{equation}
  \vect{F}(\vect{U}) = 0.
  \label{eq:nonlinearSystemResidualForm}
\end{equation}

\subsection{Newton's Method}

Newton's method if given by
\begin{itemize}
  \item[1.] Set $\vect{U}^{[0]} = \vect{U}^{n}$
  \item[2.] For $k=1,\ldots,M$ compute
  \begin{equation}
    \vect{U}^{[k]}=\vect{U}^{[k-1]} + \delta\vect{U}^{[k-1]}
  \end{equation}
  \item[3.] Set $\vect{U}^{n+1}=\vect{U}^{[M]}$,
\end{itemize}
where
\begin{equation}
  \delta\vect{U}^{[k]} = - \vect{J}(\vect{U}^{[k]})^{-1}\,\vect{F}(\vect{U}^{[k]}), \quad\text{with}\quad \vect{J}(\vect{U}^{[k]})=\vect{F}'(\vect{U}^{[k]}),
\end{equation}
and where the number of iterations $M$ is given by suitable convergence criteria.

From Eq.~\eqref{eq:electronFractionDiscrete}, the elements of the Jacobian matrix are
\begin{align}
  \pderiv{F_{Y_{e}}}{Y_{e}}
  &=\f{1}{Y_{e}^{n}}, \\
  \pderiv{F_{Y_{e}}}{\epsilon}
  &=0, \\
  \pderiv{F_{Y_{e}}}{\mathcal{J}_{j}}
  &=\f{\hat{w}_{j}^{(2)}}{S_{Y_{e}}}, \\
  \pderiv{F_{Y_{e}}}{\bar{\mathcal{J}}_{j}}
  &=-\f{\hat{w}_{j}^{(2)}}{S_{Y_{e}}}.  
\end{align}
From Eq.~\eqref{eq:internalEnergyDiscrete}, the elements of the Jacobian matrix are
\begin{align}
  \pderiv{F_{\epsilon}}{Y_{e}}
  &=0, \\
  \pderiv{F_{\epsilon}}{\epsilon}
  &=\f{1}{\epsilon^{n}}, \\
  \pderiv{F_{\epsilon}}{\mathcal{J}_{j}}
  &=\f{\hat{w}_{j}^{(3)}}{S_{\epsilon}}, \\
  \pderiv{F_{\epsilon}}{\bar{\mathcal{J}}_{j}}
  &=\f{\hat{w}_{j}^{(3)}}{S_{\epsilon}}
\end{align}
From Eq.~\eqref{eq:neutrinoDiscreteResidualForm}, the elements of the Jacobian are
\begin{align}
  \pderiv{\mathcal{F}_{i}}{Y_{e}}
  &=\dt\,\Big(\,\pderiv{\big[\chi\big]_{i}}{Y_{e}}\,\mathcal{J}_{i}-\pderiv{\big[\eta\big]_{i}}{Y_{e}}\,\Big), \\
  \pderiv{\mathcal{F}_{i}}{\epsilon}
  &=\dt\,\Big(\,\pderiv{\big[\chi\big]_{i}}{\epsilon}\,\mathcal{J}_{i}-\pderiv{\big[\eta\big]_{i}}{\epsilon}\,\Big), \\
  \pderiv{\mathcal{F}_{i}}{\mathcal{J}_{j}}
  &=\big(\,1+\dt\,\big[\chi\big]_{i}\,\big)\,\delta_{ij}
  +\dt\,\Big(\,\pderiv{\big[\chi\big]_{i}}{\mathcal{J}_{j}}\,\mathcal{J}_{i}-\pderiv{\big[\eta\big]_{i}}{\mathcal{J}_{j}}\,\Big), \\
  \pderiv{\mathcal{F}_{i}}{\bar{\mathcal{J}}_{j}}
  &=\dt\,\Big(\,\pderiv{\big[\chi\big]_{i}}{\bar{\mathcal{J}}_{j}}\,\mathcal{J}_{i}-\pderiv{\big[\eta\big]_{i}}{\bar{\mathcal{J}}_{j}}\,\Big).  
\end{align}
The derivatives of neutrino emissivities and opacities with respect to $Y_{e}$ are
\begin{align}
  \pderiv{\big[\eta\big]_{i}}{Y_{e}}
  &=\big[\chi_{\EC}\big]_{i}^{n}\,\pderiv{\mathcal{J}_{0,i}}{Y_{e}} \nonumber \\
  &\hspace{12pt}
  +\sum_{q}w_{q}^{(2)}\,\pderiv{\big[\Phi_{\NES,0}^{\In}\big]_{qi}}{Y_{e}}\,\mathcal{J}_{q}^{n+1}
  +\sum_{q}w_{q}^{(2)}\,\pderiv{\big[\Phi_{\TP,0}^{\PROD}\big]_{qi}}{Y_{e}}\,\big(1-\bar{\mathcal{J}}_{q}^{n+1}\big) , \\
  \pderiv{\big[\chi\big]_{i}}{Y_{e}}
  &=\sum_{q}w_{q}^{(2)}\,\Big\{\,\pderiv{\big[\Phi_{\NES,0}^{\In}\big]_{qi}}{Y_{e}}\,\mathcal{J}_{q}^{n+1}+\pderiv{\big[\Phi_{\NES,0}^{\Out}\big]_{qi}}{Y_{e}}\,\big(1-\mathcal{J}_{q}^{n+1}\big)\,\Big\} \nonumber \\
  &\hspace{12pt}
  +\sum_{q}w_{q}^{(2)}\Big\{\,\pderiv{\big[\Phi_{\TP,0}^{\PROD}\big]_{qi}}{Y_{e}}\,\big(1-\bar{\mathcal{J}}_{q}^{n+1}\big)+\pderiv{\big[\Phi_{\TP,0}^{\ANN}\big]_{qi}}{Y_{e}}\,\bar{\mathcal{J}}_{q}^{n+1}\,\Big\}.  
\end{align}
The derivatives of neutrino emissivities and opacities with respect to $\epsilon$ are
\begin{align}
  \pderiv{\big[\eta\big]_{i}}{\epsilon}
  &=\big[\chi_{\EC}\big]_{i}^{n}\,\pderiv{\mathcal{J}_{0,i}}{\epsilon} \nonumber \\
  &\hspace{12pt}
  +\sum_{q}w_{q}^{(2)}\,\pderiv{\big[\Phi_{\NES,0}^{\In}\big]_{qi}}{\epsilon}\,\mathcal{J}_{q}^{n+1}
  +\sum_{q}w_{q}^{(2)}\,\pderiv{\big[\Phi_{\TP,0}^{\PROD}\big]_{qi}}{\epsilon}\,\big(1-\bar{\mathcal{J}}_{q}^{n+1}\big) , \\
  \pderiv{\big[\chi\big]_{i}}{\epsilon}
  &=\sum_{q}w_{q}^{(2)}\,\Big\{\,\pderiv{\big[\Phi_{\NES,0}^{\In}\big]_{qi}}{\epsilon}\,\mathcal{J}_{q}^{n+1}+\pderiv{\big[\Phi_{\NES,0}^{\Out}\big]_{qi}}{\epsilon}\,\big(1-\mathcal{J}_{q}^{n+1}\big)\,\Big\} \nonumber \\
  &\hspace{12pt}
  +\sum_{q}w_{q}^{(2)}\Big\{\,\pderiv{\big[\Phi_{\TP,0}^{\PROD}\big]_{qi}}{\epsilon}\,\big(1-\bar{\mathcal{J}}_{q}^{n+1}\big)+\pderiv{\big[\Phi_{\TP,0}^{\ANN}\big]_{qi}}{\epsilon}\,\bar{\mathcal{J}}_{q}^{n+1}\,\Big\}.  
\end{align}
The derivatives of neutrino emissivities and opacities with respect to $\mathcal{J}_{j}$ are
\begin{align}
  \pderiv{\big[\eta\big]_{i}}{\mathcal{J}_{j}}
  &=w^{(2)}_{j}\,\big[\Phi_{\NES,0}^{\In}\big]_{ji}, \\
  \pderiv{\big[\chi\big]_{i}}{\mathcal{J}_{j}}
  &=w^{(2)}_{j}\,\Big(\,\big[\Phi_{\NES,0}^{\In}\big]_{ji}-\big[\Phi_{\NES,0}^{\Out}\big]_{ji}\,\Big).  
\end{align}
The derivatives of neutrino emissivities and opacities with respect to $\bar{\mathcal{J}}_{j}$ are
\begin{align}
  \pderiv{\big[\eta\big]_{i}}{\bar{\mathcal{J}}_{j}}
  &=-w_{j}^{(2)}\,\big[\Phi_{\TP,0}^{\PROD}\big]_{ji} , \\
  \pderiv{\big[\chi\big]_{i}}{\bar{\mathcal{J}}_{j}}
  &=w_{j}^{(2)}\,\big(\,\big[\Phi_{\TP,0}^{\ANN}\big]_{ji}-\big[\Phi_{\TP,0}^{\PROD}\big]_{ji}\,\big) .
\end{align}

@@

\appendix

\section{Kernel Derivatives}

NES and pair kernels are tabulated in terms of temperature $T$ and degeneracy parameter $\eta=\mu_{e}/kT$
\begin{equation}
  \Phi=\Phi(T,\eta).  
\end{equation}
For solvers based on Newton's method, we need derivatives with respect to internal energy $\epsilon$ and electron fraction $Y_{e}$.  
\begin{align}
  d\Phi 
  &= \Big(\pderiv{\Phi}{T}\Big)_{\eta}\,dT+\Big(\pderiv{\Phi}{\eta}\Big)_{T}\,d\eta \nonumber \\
  &=\f{1}{kT}\Big(\pderiv{\Phi}{\eta}\Big)_{T}\,\Big(\pderiv{\mu_{e}}{\rho}\Big)_{T,Y_{e}}\,d\rho
  +\Big\{\,\Big(\pderiv{\Phi}{T}\Big)_{\eta} + \Big(\pderiv{\Phi}{\eta}\Big)_{T}\,\Big[\,\f{1}{kT}\Big(\pderiv{\mu_{e}}{T}\Big)_{\rho,Y_{e}}-\f{\eta}{T}\,\Big]\,\Big\}\,dT \nonumber \\
  &\hspace{12pt}
  +\f{1}{kT}\Big(\pderiv{\Phi}{\eta}\Big)_{T}\,\Big(\pderiv{\mu_{e}}{Y_{e}}\Big)_{\rho,T}\,dY_{e},
  \label{eq:variation_T_Eta}
\end{align}
where we used
\begin{align}
  d\eta &= \Big(\pderiv{\eta}{\rho}\Big)_{T,Y_{e}}\,d\rho+\Big(\pderiv{\eta}{T}\Big)_{\rho,Y_{e}}\,dT+\Big(\pderiv{\eta}{Y_{e}}\Big)_{\rho,T}\,dY_{e} \nonumber \\
  &=\f{1}{kT}\Big(\pderiv{\mu_{e}}{\rho}\Big)_{T,Y_{e}}\,d\rho 
  + \Big[\,\f{1}{kT}\Big(\pderiv{\mu_{e}}{T}\Big)_{\rho,Y_{e}}-\f{\eta}{T}\,\Big]\,dT + \f{1}{kT}\Big(\pderiv{\mu_{e}}{Y_{e}}\Big)_{\rho,T}\,dY_{e}.  
\end{align}

\begin{align}
  d\Phi
  &=\Big(\pderiv{\Phi}{\epsilon}\Big)_{Y_{e}}\,d\epsilon+\Big(\pderiv{\Phi}{Y_{e}}\Big)_{\epsilon}\,dY_{e} \nonumber \\
  &=\Big(\pderiv{\Phi}{\epsilon}\Big)_{Y_{e}}\,\Big(\pderiv{\epsilon}{\rho}\Big)_{T,Y_{e}}\,d\rho
  +\Big(\pderiv{\Phi}{\epsilon}\Big)_{Y_{e}}\,\Big(\pderiv{\epsilon}{T}\Big)_{\rho,Y_{e}}\,dT \nonumber \\
  &\hspace{12pt}
  +\Big[\,\Big(\pderiv{\Phi}{\epsilon}\Big)_{Y_{e}}\,\Big(\pderiv{\epsilon}{Y_{e}}\Big)_{\rho,T}+\Big(\pderiv{\Phi}{Y_{e}}\Big)_{\epsilon}\,\Big]\,dY_{e},
  \label{eq:variation_Epsilon_Ye}
\end{align}
where we used
\begin{equation}
  d\epsilon
  =\Big(\pderiv{\epsilon}{\rho}\Big)_{T,Y_{e}}\,d\rho+\Big(\pderiv{\epsilon}{T}\Big)_{\rho,Y_{e}}\,dT+\Big(\pderiv{\epsilon}{Y_{e}}\Big)_{\rho,T}\,dY_{e}
\end{equation}
Comparing Eq.~\eqref{eq:variation_T_Eta} and Eq.~\eqref{eq:variation_Epsilon_Ye}, we have
\begin{align}
  \Big(\pderiv{\Phi}{T}\Big)_{\eta} + \Big[\,\f{1}{kT}\Big(\pderiv{\mu_{e}}{T}\Big)_{\rho,Y_{e}}-\f{\eta}{T}\,\Big]\,\Big(\pderiv{\Phi}{\eta}\Big)_{T}
  &=\Big(\pderiv{\epsilon}{T}\Big)_{\rho,Y_{e}}\,\Big(\pderiv{\Phi}{\epsilon}\Big)_{Y_{e}}, \\
  \f{1}{kT}\Big(\pderiv{\mu_{e}}{Y_{e}}\Big)_{\rho,T}\,\Big(\pderiv{\Phi}{\eta}\Big)_{T}
  &=\Big(\pderiv{\epsilon}{Y_{e}}\Big)_{\rho,T}\,\Big(\pderiv{\Phi}{\epsilon}\Big)_{Y_{e}}+\Big(\pderiv{\Phi}{Y_{e}}\Big)_{\epsilon}
\end{align}
Solving for $(\partial\Phi/\partial\epsilon)_{Y_{e}}$ and $(\partial\Phi/\partial Y_{e})_{\epsilon}$ we obtain
\begin{align}
  \Big(\pderiv{\Phi}{\epsilon}\Big)_{Y_{e}}
  &=\Big[\,\f{1}{kT}\Big(\pderiv{\mu_{e}}{T}\Big)_{\rho,Y_{e}}-\f{\eta}{T}\,\Big]/\Big(\pderiv{\epsilon}{T}\Big)_{\rho,Y_{e}}\,\Big(\pderiv{\Phi}{\eta}\Big)_{T} + \Big(\pderiv{\epsilon}{T}\Big)_{\rho,Y_{e}}^{-1}\,\Big(\pderiv{\Phi}{T}\Big)_{\eta}, \\
  \Big(\pderiv{\Phi}{Y_{e}}\Big)_{\epsilon}
  &=\Big\{\,\f{1}{kT}\Big(\pderiv{\mu_{e}}{Y_{e}}\Big)_{\rho,T} - \Big(\pderiv{\epsilon}{Y_{e}}\Big)_{\rho,T}\,\Big[\,\f{1}{kT}\Big(\pderiv{\mu_{e}}{T}\Big)_{\rho,Y_{e}}-\f{\eta}{T}\,\Big]/\Big(\pderiv{\epsilon}{T}\Big)_{\rho,Y_{e}}\,\Big\}\,\Big(\pderiv{\Phi}{\eta}\Big)_{T} \nonumber \\
  &\hspace{12pt}
  - \Big\{\,\Big(\pderiv{\epsilon}{Y_{e}}\Big)_{\rho,T}/\Big(\pderiv{\epsilon}{T}\Big)_{\rho,Y_{e}}\,\Big\}\,\Big(\pderiv{\Phi}{T}\Big)_{\eta}.
\end{align}
